Multi-Instance Methods for Partially Supervised Image Segmentation

### 2013-09-16 partial_supervision/paper.tex
rename MIK-Instance to MIK-I for tables ^^ also some minor fixes.

% TODO contrast with Alex

%The task of multi-class segmentation is often split up in a segmentation
%and a recognition part. Random forest methods often classify each pixel
%separately and segment using predicted classes~\citep{schroff2008object} while SVM-based methods often work on an
%over-segmentation of the image, called superpixels~\citep{ladicky2009associative,gonfaus2010harmony}. Superpixels avoid the
%computational burden of classifying each pixel separately, but have
%two drawbacks:
%\begin{enumerate}
    %\item A single superpixel does not provide enough context for classification~\citep{fulkerson2009class}.
    %\item Segment boundaries are decided on the lowest level by generating
        %superpixels. This decision cannot be corrected afterwards~\citep{hanbury2008superpixels}.
%\end{enumerate}
In our approach, we work with a set of candidate segments, generated using
constrained parametric min-cuts~\citep{carreira2010constrained}.  For each
image, these segments are a set of overlapping, object-like regions, which
serve as candidates for object locations.

@@

%\subsection{Multi Instance Multi Label (MIML) Methods}
%Zhou and Zhang~\citet{zhou2007multi} were some of the first to tackle the problem of multi-instance multi-label tasks.
%They proposed two methods, MIMLBoost and MIMLSvm. MIMLBoost transforms a MIML task into
%a multi-instance learning task by reformulating the problem of \eqref{miml} as
%\begin{equation} f_{MIL} \colon 2^\mathcal{X}\times \mathcal{Y} \rightarrow \{ -1, +1\}.\end{equation}
%This corresponds to learning a single multi-instance classifier per class.
%MIMLSvm, on the other hand, transforms a MIML task into a multi-label task by constructing a mapping
%\begin{equation} \phi \colon 2^\mathcal{X} \rightarrow \mathcal{Z} \end{equation}
%into a new instance space $\mathcal{Z}$ by medioid clustering.
%Our approach fits well into the category of MIMLBoost: we also transform the MIML task
%into a multi-instance task that we then solve using MI-kernels.
%While our problem formulation is similar to Zhou and Zhang, we aim not only at image classification but
%also at object segmentation.

@@

While using MIK has many advantages, it produces only an instance-level
classifier. We propose to transform a bag-level classifier $f_{MI}$ as given by
the SVM and \eqref{mi_kernel} into an instance-level classifier by setting
$f_{I}(x):=f_{MI}(\{x\})$, in other words, by considering each instance as its own
bag. 
We extend the MIK SVM to the multi-class case using the one-vs-rest approach,
learning one SVM for each class.

@@

\subsection{Instance-Level Predictions using Multi-Instance Kernels}

To assess the validity of instance-level predictions using multi-instance
kernels, we transform $f_{I}$ back to an instance-level classifier, using the
multi-instance learning assumption (\eqref{multi_instance}). We refer to these
instance-based MIK predictions as MIK-instance. In all experiments, the
parameters of the MI-Kernel and SVM are adjusted using MIK and then used with
both MIK and MIK-instance.  This facilitates very fast parameter scans since
MIK is very efficient to compute. Note that we cannot adjust parameters using
instance prediction error, as we assume no instance labels to be known.

\begin{table}
    \centering
    \begin{tabularx}{\linewidth}{@{\extracolsep{\fill}}lccccccc}
    \toprule
        & SVM-SVR & EMDD & mi-SVM & MI-SVM & MICA & MIK & MIK-instance \\
    \cmidrule(rl){2-8}
    Musk1 & 87.9 &84.9 &  87.4 &  77.9     & 84.3 & 88.0& 88.0 \\
    Musk2 & 85.4 &84.8 &  83.6 &  84.3     & 90.5 & 89.3& 85.2 \\
    \bottomrule
    \end{tabularx}
    \caption{Bag level performance of various MIL algorithms on the standard Musk
    datasets. All but MIK provide instance-level labeling. \tablabel{muskacc}}

\end{table}
We compared the performance of MIK, MIK-instance and state-of-the-art MI
methods on the Musk benchmark datasets~\citep{dietterich1997solving}, see
\Tabref{muskacc}. Results were obtained using 10-fold cross-validation. While
the computational complexity of MIK-instance is very low compared to the other
methods, it achieves competitive results.  Using instance-level labels results
in a slight loss of accuracy of MIK-instances, compared to MIK\@. This small
degradation of performance is quite surprising, since the model was not trained
to provide any instance-level labels.

For multi-class image segmentation, it is beneficial to have a low witness
rate, that is only a few instances are assumed to be positive in a positive
bag. Since an object might not be very prominent in an image, only a fraction
of segments might correspond to the object.
\Tabref{musk-witness} compares the witness rates of MIK-instance,
miSVM~\citep{andrews2003support} and SVR-SVM~\citep{liconvex2010} on the Musk
datasets. MIK-instance is able to achieve similar accuracy with much less
witnesses than the other methods.  Note that Musk1 consists of very small bags
while Musk2 contains significantly larger bags, more similar to the
image/segment setup.

\begin{table}
    \centering
    \begin{tabularx}{\linewidth}{@{\extracolsep{\fill}}lcccc}
    \toprule
    & \multicolumn2c{Musk1}  & \multicolumn2c{Musk2}  \\
                &accuracy & witness-rate & accuracy & witness-rate  \\
    \cmidrule(rl){2-3}
    \cmidrule(rl){4-5}
    mi-SVM      & 87.4          & 100\%               &  83.6          & 83.9\%\\
    SVM-SVR     & 87.9          & 100\%               &  85.4          & 89.5\%\\
    MIK-instance& 88.0          & 99\%                &  85.2          & 62.3\%\\
    \bottomrule
    \end{tabularx}
    \caption{MIL algorithms and the empirical witness rates of the
        classifiers.  \tablabel{musk-witness}}
\end{table}

@@

We adjusted parameters on a hold-out validation set using bag-level information
and used the training and test sets as given by the dataset.  It is
straight-forward to extend the binary MIK method to the multi-class setting
using a one-vs-all strategy.
We train one multiple kernel learning (MKL) SVM per class using MIK and predict
class labels on segment level using MIK-instance. If at least one SVM
classifies a segment as positive, it is associated with the most confident
class. Otherwise, it is assigned ``background'' or no class.
This yields a classification of each segment into one of four classes: car,
bike, person, or background. We merge segments into pixel-level class labels as
described in \Secref{combiningsegments}.

@@

Per-class pixel accuracies are reported in \Tabref{tab:graz}; some qualitative
results are shown in \figref{results}. The overall accuracy on images labels,
which is the task that was actually trained, is $90\%$.  While the performance
of our multiple-instance based approach is far from current methods that use
pixel-level annotations, whose pixel-level accuracy is around
$70\%$~\citep{fulkerson2009class,schulz2011} on pixel-level, it can serve as a
baseline for research on weakly supervised methods for object-class segmentation.

>>>>

Recently, several approaches have been proposed for semi-supervised semantic segmentation.
While these are close to our work, there are several important distinctions.
We address the task of object-class segmentation, that concerns object categories, while
semantic segmentation approaches often focus on "stuff" categories like "sky" and "road" that
are more easily detected using simple texture features.
Additionally, in contrast to \citet{vezhnevets2011weakly}, our approach is in
principle applicable to large datasets, the regime where weak annotation is
most useful.

In our approach, we work with a set of candidate segments, generated using
constrained parametric min-cuts~\citep{carreira2010constrained}.  For each
image, these segments are a set of overlapping, object-like regions, which
serve as candidates for object locations.

@@

%\subsection{Multi Instance Multi Label (MIML) Methods}
%Zhou and Zhang~\citet{zhou2007multi} were some of the first to tackle the problem of multi-instance multi-label tasks.
%They proposed two methods, MIMLBoost and MIMLSvm. MIMLBoost transforms a MIML task into
%a multi-instance learning task by reformulating the problem of \Eqref{miml} as
%\begin{equation} f_{MIL} \colon 2^\mathcal{X}\times \mathcal{Y} \rightarrow \{ -1, +1\}.\end{equation}
%This corresponds to learning a single multi-instance classifier per class.
%MIMLSvm, on the other hand, transforms a MIML task into a multi-label task by constructing a mapping
%\begin{equation} \phi \colon 2^\mathcal{X} \rightarrow \mathcal{Z} \end{equation}
%into a new instance space $\mathcal{Z}$ by medioid clustering.
%Our approach fits well into the category of MIMLBoost: we also transform the MIML task
%into a multi-instance task that we then solve using MI-kernels.
%While our problem formulation is similar to Zhou and Zhang, we aim not only at image classification but
%also at object segmentation.

@@

While using MIK has many advantages, it produces only an instance-level
classifier. We propose to transform a bag-level classifier $f_{MI}$ as given by
the SVM and \Eqref{mi_kernel} into an instance-level classifier by setting
$f_{I}(x):=f_{MI}(\{x\})$, in other words, by considering each instance as its own
bag. 
We extend the MIK SVM to the multi-class case using the one-vs-rest approach,
learning one SVM for each class.

@@

\subsection{Instance-Level Predictions using MI-Kernel}

To assess the validity of instance-level predictions using multi-instance
kernels, we transform $f_{I}$ back to an instance-level classifier, using the
multi-instance learning assumption (\Eqref{multi_instance}). We refer to these
instance-based MIK predictions as MIK-I. In all experiments, the
parameters of the MI-Kernel and SVM are adjusted using MIK and then used with
both MIK and MIK-I.  This facilitates very fast parameter scans since
MIK is very efficient to compute. Note that we cannot adjust parameters using
instance prediction error, as we assume no instance labels to be known.

\begin{table}
    \centering
    \begin{tabularx}{\linewidth}{@{\extracolsep{\fill}}lccccccc}
    \toprule
        & SVM-SVR & EMDD & mi-SVM & MI-SVM & MICA & MIK & MIK-I \\
    \cmidrule(rl){2-8}
    Musk1 & 87.9 &84.9 &  87.4 &  77.9     & 84.3 & 88.0& 88.0 \\
    Musk2 & 85.4 &84.8 &  83.6 &  84.3     & 90.5 & 89.3& 85.2 \\
    \bottomrule
    \end{tabularx}
    \caption{Bag level performance of various MIL algorithms on the standard Musk
    datasets. All but MIK provide instance-level labeling.\tablabel{muskacc}}

\end{table}
We compared the performance of MIK, MIK-I and state-of-the-art MI
methods on the Musk benchmark datasets~\citep{dietterich1997solving}, see
\Tabref{muskacc}. Results were obtained using 10-fold cross-validation. While
the computational complexity of MIK-I is very low compared to the other
methods, it achieves competitive results.  Using instance-level labels results
in a slight loss of accuracy of MIK-I, compared to MIK\@. This small
degradation of performance is quite surprising, since the model was not trained
to provide any instance-level labels.

For multi-class image segmentation, it is beneficial to have a low witness
rate, that is only a few instances are assumed to be positive in a positive
bag. Since an object might not be very prominent in an image, only a fraction
of segments might correspond to the object.
\Tabref{musk-witness} compares the witness rates of MIK-I,
miSVM~\citep{andrews2003support} and SVR-SVM~\citep{liconvex2010} on the Musk
datasets. MIK-I is able to achieve similar accuracy with much less
witnesses than the other methods.  Note that Musk1 consists of very small bags
while Musk2 contains significantly larger bags, more similar to the
image/segment setup.

\begin{table}
    \centering
    \begin{tabularx}{\linewidth}{@{\extracolsep{\fill}}lcccc}
    \toprule
    & \multicolumn2c{Musk1}  & \multicolumn2c{Musk2}  \\
                &accuracy & witness-rate & accuracy & witness-rate  \\
    \cmidrule(rl){2-3}
    \cmidrule(rl){4-5}
    mi-SVM      & 87.4          & 100\%               &  83.6          & 83.9\%\\
    SVM-SVR     & 87.9          & 100\%               &  85.4          & 89.5\%\\
    MIK-I& 88.0          & 99\%                &  85.2          & 62.3\%\\
    \bottomrule
    \end{tabularx}
    \caption{MIL algorithms and the empirical witness rates of the
        classifiers.  \tablabel{musk-witness}}
\end{table}

@@

We adjusted parameters on a hold-out validation set using bag-level information
and used the training and test sets as given by the dataset.  It is
straight-forward to extend the binary MIK method to the multi-class setting
using a one-vs-all strategy.
We train one multiple kernel learning (MKL) SVM per class using MIK and predict
class labels on segment level using MIK-I. If at least one SVM
classifies a segment as positive, it is associated with the most confident
class. Otherwise, it is assigned ``background'' or no class.
This yields a classification of each segment into one of four classes: car,
bike, person, or background. We merge segments into pixel-level class labels as
described in \Secref{combiningsegments}.

@@

Per-class pixel accuracies are reported in \Tabref{tab:graz}; some qualitative
results are shown in \Figref{results}. The overall accuracy on images labels,
which is the task that was actually trained, is $90\%$.  While the performance
of our multiple-instance based approach is far from current methods that use
pixel-level annotations, whose pixel-level accuracy is around
$70\%$~\citep{fulkerson2009class,schulz2011} on pixel-level, it can serve as a
baseline for research on weakly supervised methods for object-class segmentation.
